\chapter{Installation}

\section{Downloading the Repository}
The repository is available on \textbf{github} through URL : \url{https://github.com/cs207-group11/cs207-FinalProject.git}. To download :
\begin{itemize}
    \item Below 'Contributors', a green button to 'Clone or download' the repository is present.
    \item Use the 'Download ZIP' option to begin the download.
    \item Once completed, the ZIP file will be present in 'Downloads' folder of your computer (default).
    \item Unzip this file into a directory of your choice using WinRAR or equivalent software.
    \item Now, follow instructions in Chapter 3 : Procedure for usage.
\end{itemize}
This link can also be used to clone the repository to your machine through the Mac Terminal/GitBash.
\begin{itemize}
    \item Enter directory of your choice using appropriate 'cd' commands.
\end{itemize} 
\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=5, innerleftmargin=10, innertopmargin=5,innerbottommargin=5, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
$ git clone <repository URL mentioned above>
$ cd cs207-FinalProject
\end{lstlisting}
\end{mdframed} 
You are now in the repository and can follow Chapter 3 : Procedure instructions for usage.
\newpage
\section{Test Suite}
The testing is an important phase of development of any project and a variety of tests are necessary before the design is released. We have defined a test module called 'test\_chemkin.py' within this repository.
\subsection{Types of Errors}
\begin{itemize}
    \item ValueError: check the user's input values for incorrect variable or parameter definitions. Ex : Negative Temperature/Molar Concentration. 
    \item AttributeError: check if the execution of the program is in the right order. Ex : If the Rate Coefficient has been calculated before attempting to calculate the Reaction Rate.
    \item TypeError : checks if the entered parameters are real and belong to a certain data type. Ex : Modified Arrhenius factor 'b' is real and a float.
    \item NotImplementedError : checks if a certain feature/function has been designed or is due for implementation.
\end{itemize}
The above types of errors are tested for within the module 'test\_chemkin.py' for erroneous input to test the robustness of the 'chemkin.py' module.
\subsection{Running the Test Suite}
The test suite can be run from the \textbf{Terminal} to view the results of the testing.
\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=5, innerleftmargin=10, innertopmargin=5,innerbottommargin=5, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
#Enter repository directory
>>> python -m pytest test_chemkin.py
\end{lstlisting}
\end{mdframed} 